---
title: "Master's Thesis: Mainstream-Party Responses to the Populist Radical Right"
subtitle: "Comparative analysis — CHES trend data, 1999–2024"
author: "Carlos Morán García-Quijada"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
  pdf_document:
    toc: true
    toc_depth: 2
---

Code accompanying the master's thesis. Data: Chapel Hill Expert Survey trend
file 1999-2024 (Rovny et al., 2025). Run the chunks in order.

```{r setup, include=FALSE}
# --- KNITR DEFAULTS ---
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = TRUE)
```

# 1. Setup and data

## 1.1 Packages and reproducibility

```{r packages}
# --- PACKAGES ---
# fixest: TWFE benchmark and Sun-Abraham event study
# did: Callaway-Sant'Anna estimator (main specification)
# flextable / stargazer / kableExtra: tables exported to the thesis
library(tidyverse)
library(fixest)
library(did)

library(knitr)
library(kableExtra)
library(flextable)
library(stargazer)
library(patchwork)
```

```{r table-formatting-function}
# --- SHARED TABLE FORMATTER ---
# One helper so every ATT table in the thesis shares identical styling
# (Times New Roman 10 pt, booktabs rules, note + source footer).
format_att_table <- function(data, caption) {
  
  data |>
    mutate(
      `ATT (SE)` = sprintf("%.3f (%.3f)", att, se),
      `95% CI` = sprintf("[%.3f, %.3f]", lower, upper)
    ) |>
    select(sample, outcome, `ATT (SE)`, `95% CI`) |>
    kable(
      format = "html",
      caption = caption,
      col.names = c("Sample", "Outcome", "ATT (SE)", "95% CI"),
      align = c("l", "l", "c", "c"),
      escape = FALSE
    ) |>
    kable_styling(
      full_width = FALSE,
      position = "center",
      bootstrap_options = c("striped", "hover", "condensed")
    )
}
```

## 1.2 Load CHES trend file

```{r load-data}
# --- LOAD CHES TREND FILE (1999-2024) ---
# Raw trend file as distributed by Rovny et al. (2025); nothing is edited by hand.
ches_raw <- read_csv(
  "1999-2024_CHES_dataset_meansV2.csv",
  show_col_types = FALSE
)

ches <- ches_raw
```

## 1.3 Labels and technical checks

```{r labels}
# --- COUNTRY AND PARTY-FAMILY LABELS ---
# CHES ships numeric codes; readable labels are attached once here so every
# downstream table and figure inherits them.
country_labels <- c(
  `1` = "Belgium",
  `2` = "Denmark",
  `3` = "Germany",
  `4` = "Greece",
  `5` = "Spain",
  `6` = "France",
  `7` = "Ireland",
  `8` = "Italy",
  `10` = "Netherlands",
  `11` = "United Kingdom",
  `12` = "Portugal",
  `13` = "Austria",
  `14` = "Finland",
  `16` = "Sweden",
  `20` = "Bulgaria",
  `21` = "Czech Republic",
  `22` = "Estonia",
  `23` = "Hungary",
  `24` = "Latvia",
  `25` = "Lithuania",
  `26` = "Poland",
  `27` = "Romania",
  `28` = "Slovakia",
  `29` = "Slovenia",
  `31` = "Croatia",
  `37` = "Malta",
  `38` = "Luxembourg",
  `40` = "Cyprus"
)

family_labels <- c(
  `1` = "Radical right",
  `2` = "Conservative",
  `3` = "Liberal",
  `4` = "Christian-democratic",
  `5` = "Socialist",
  `6` = "Radical left",
  `7` = "Green",
  `8` = "Regionalist",
  `9` = "No family",
  `10` = "Confessional",
  `11` = "Agrarian/Centre"
)

ches <- ches |>
  mutate(
    country_name = recode(as.character(country), !!!country_labels),
    family_name = recode(as.character(family), !!!family_labels),
    region_name = case_when(
      eastwest == 1 ~ "Western Europe",
      eastwest == 0 ~ "Central/Eastern Europe",
      TRUE ~ NA_character_
    )
  )
```

```{r technical-checks, include=FALSE}
# --- SANITY CHECKS (not reported) ---
# include=FALSE keeps these out of the knitted document.
stopifnot(!anyDuplicated(ches[c("party_id", "year")]))

technical_checks <- list(
  data_dimensions = dim(ches),
  waves = ches |> distinct(year) |> arrange(year),
  party_year_duplicates = ches |> count(party_id, year) |> filter(n > 1),
  party_multiple_countries = ches |> distinct(party_id, country) |>
    count(party_id) |> filter(n > 1),
  vote_seat_ranges = ches |>
    summarise(
      vote_min = min(vote, na.rm = TRUE),
      vote_max = max(vote, na.rm = TRUE),
      seat_min = min(seat, na.rm = TRUE),
      seat_max = max(seat, na.rm = TRUE)
    )
)
```

## 1.4 Outcome coverage

```{r outcome-coverage}
# --- ITEM COVERAGE BY WAVE ---
# Documents why immigration models start in 2006: the immigration items enter
# CHES in that wave, and immigration salience is missing in 2014.
outcomes_core <- c(
  "galtan",
  "immigrate_policy",
  "immigrate_salience",
  "multiculturalism",
  "lrecon"
)

outcome_coverage <- ches |>
  group_by(year) |>
  summarise(
    across(
      all_of(outcomes_core),
      ~ round(mean(!is.na(.x)) * 100, 1)
    ),
    .groups = "drop"
  )

outcome_coverage
```

# 2. Treatment and analytical sample

## 2.1 Radical-right parties and treatment definition

```{r radical-right-identification}
# --- TREATMENT DEFINITION ---
# A country is treated in the first wave in which a CHES-classified radical
# right party (family == 1) holds at least 5% of national parliamentary seats.
# Seats rather than votes: under a vote threshold France would already be
# treated in 1999, losing its entire pre-treatment window.
ches <- ches |>
  mutate(is_prrp_ches = family == 1)

prrp_seat_summary <- ches |>
  filter(
    is_prrp_ches,
    eastwest == 1
  ) |>
  group_by(country, country_name, year) |>
  summarise(
    max_prrp_seat = if (all(is.na(seat))) {
  NA_real_
} else {
  max(seat, na.rm = TRUE)
},
    prrp_parties = paste(
      party[!is.na(seat) & seat >= 5],
      collapse = " / "
    ),
    .groups = "drop"
  ) |>
  mutate(breakthrough = !is.na(max_prrp_seat) & max_prrp_seat >= 5)

treatment_cohorts <- prrp_seat_summary |>
  filter(breakthrough) |>
  group_by(country, country_name) |>
  summarise(
    treatment_year = min(year),
    .groups = "drop"
  ) |>
  arrange(treatment_year, country)

treatment_cohorts
```

## 2.2 Country-level cohorts

```{r country-cohorts}
# --- COUNTRY COHORTS ---
# Countries already treated in the first observed wave have no pre-treatment
# period and are therefore dropped from the estimation sample.
western_country_cohorts <- ches |>
  filter(eastwest == 1) |>
  distinct(country, country_name) |>
  left_join(
    treatment_cohorts,
    by = c("country", "country_name")
  ) |>
  mutate(
    first_wave = min(ches$year),
    treatment_status = case_when(
      is.na(treatment_year) ~ "Never treated",
      treatment_year == first_wave ~ "Treated from first wave",
      TRUE ~ "Treated during observation window"
    )
  ) |>
  select(
    country,
    country_name,
    treatment_year,
    treatment_status
  ) |>
  arrange(treatment_year, country)

usable_country_cohorts <- western_country_cohorts |>
  filter(treatment_status != "Treated from first wave")

usable_country_cohorts
```

### 2.2.1 Country cohort table

```{r cohort-table}
# --- COHORT TABLE DATA (Table 2.1 in the thesis) ---
cohort_table <- western_country_cohorts |>
  mutate(
    empirical_role = case_when(
      treatment_status == "Never treated" ~
        "Never-treated (pure control group)",
      treatment_status == "Treated from first wave" ~
        "Treated from first wave (excluded)",
      TRUE ~ "Treatment cohort (staggered adoption)"
    )
  ) |>
  group_by(empirical_role, treatment_year) |>
  summarise(
    countries = paste(sort(country_name), collapse = ", "),
    .groups = "drop"
  ) |>
  arrange(
    factor(
      empirical_role,
      levels = c(
        "Treatment cohort (staggered adoption)",
        "Treated from first wave (excluded)",
        "Never-treated (pure control group)"
      )
    ),
    treatment_year
  )

cohort_table
```

```{r table-cohorts}
# --- FORMAT AND EXPORT TABLE 2.1 ---
table_cohorts <- cohort_table |>
  mutate(
    wave = if_else(is.na(treatment_year), "\u2014", as.character(treatment_year))
  ) |>
  select(empirical_role, wave, countries) |>
  flextable() |>
  set_header_labels(
    empirical_role = "Empirical role",
    wave           = "Breakthrough wave (year)",
    countries      = "Countries"
  ) |>
  merge_v(j = "empirical_role") |>                # merge the role label into a single block
  valign(j = "empirical_role", valign = "top") |>
  theme_booktabs() |>
  align(align = "left", part = "all") |>
  bold(part = "header") |>
  font(fontname = "Times New Roman", part = "all") |>
  fontsize(size = 10, part = "all") |>
  hline(i = c(5, 6), border = officer::fp_border(width = 0.5)) |>  # rule separating the three groups
  add_footer_lines(
    c(
      "Note. Western European staggered treatment cohorts are relatively small and occasionally contain a single country. Consequently, cohort-specific estimates should be interpreted with caution.",
      "Source. Own elaboration based on the 1999\u20132024 Chapel Hill Expert Survey trend file (Rovny et al., 2025)."
    )
  ) |>
  italic(part = "footer") |>
  fontsize(size = 9, part = "footer") |>
  autofit()

table_cohorts

save_as_docx(table_cohorts, path = "table_cohorts.docx")
```

## 2.3 Party-wave DiD panel

```{r did-panel}
# --- PARTY-WAVE ANALYTICAL PANEL ---
# Mainstream = non-PRRP parties from the four traditional governing families
# (CHES families 2-5: conservative, liberal, Christian-democratic, socialist).
# wave: ordinal wave index (1-7) instead of calendar year, so cohorts stay
# comparable despite irregular CHES intervals.
# g_wave = 0 flags never-treated units, as att_gt() expects.
ches <- ches |>
  mutate(wave = match(year, sort(unique(year))))

usable_country_cohorts <- usable_country_cohorts |>
  mutate(
    g_wave = case_when(
      is.na(treatment_year) ~ 0L,
      TRUE ~ match(treatment_year, sort(unique(ches$year)))
    )
  )

did_panel <- ches |>
  filter(
    eastwest == 1,
    !is_prrp_ches,
    family %in% c(2, 3, 4, 5)
  ) |>
  left_join(
    usable_country_cohorts |>
      select(
        country,
        country_name,
        treatment_year,
        treatment_status,
        g_wave
      ),
    by = c("country", "country_name")
  ) |>
  filter(!is.na(treatment_status)) |>
  select(
    party_id,
    party,
    country,
    country_name,
    year,
    wave,
    family,
    family_name,
    vote,
    seat,
    treatment_year,
    treatment_status,
    g_wave,
    galtan,
    immigrate_policy,
    immigrate_salience,
    multiculturalism,
    lrecon
  ) |>
  mutate(
    treated = case_when(
      g_wave == 0 ~ 0L,
      wave >= g_wave ~ 1L,
      TRUE ~ 0L
    )
  )

sample_overview <- did_panel |>
  summarise(
    countries = n_distinct(country),
    parties = n_distinct(party_id),
    observations = n(),
    treated_countries = n_distinct(country[g_wave > 0]),
    never_treated_countries = n_distinct(country[g_wave == 0])
  )

sample_overview
```

## 2.4 Electoral-size thresholds

```{r electoral-size-thresholds}
# --- RESTRICTED SAMPLES BY PRE-TREATMENT SIZE ---
# Mean vote share is computed only over each party's pre-treatment waves, so
# sample membership cannot be affected by the breakthrough itself.
party_size_pre <- did_panel |>
  filter(is.na(treatment_year) | year < treatment_year) |>
  group_by(country, country_name, party_id) |>
  summarise(
    party_names = paste(sort(unique(party)), collapse = " / "),
    pre_treatment_mean_vote = if_else(
      all(is.na(vote)),
      NA_real_,
      mean(vote, na.rm = TRUE)
    ),
    n_pre_treatment_waves = sum(!is.na(vote)),
    .groups = "drop"
  )

party_ids_5 <- party_size_pre |>
  filter(pre_treatment_mean_vote >= 5) |>
  pull(party_id)

party_ids_10 <- party_size_pre |>
  filter(pre_treatment_mean_vote >= 10) |>
  pull(party_id)

did_panel_5 <- did_panel |>
  filter(party_id %in% party_ids_5)

did_panel_10 <- did_panel |>
  filter(party_id %in% party_ids_10)

threshold_overview <- tibble(
  sample = c("Broad sample", "At least 5% pre-treatment vote", "At least 10% pre-treatment vote"),
  parties = c(
    n_distinct(did_panel$party_id),
    n_distinct(did_panel_5$party_id),
    n_distinct(did_panel_10$party_id)
  ),
  observations = c(nrow(did_panel), nrow(did_panel_5), nrow(did_panel_10))
)

threshold_overview
```

### 2.4.1 Nested analytical sample configurations

```{r nested-sample-config-table}
# --- NESTED SAMPLE TABLE DATA (Appendix Table A.1) ---
sample_table <- threshold_overview |>
  mutate(
    selection_criterion = case_when(
      sample == "Broad sample" ~
        "All non-PRRP mainstream parties (conservative, liberal, Christian-democratic, socialist)",
      sample == "At least 5% pre-treatment vote" ~
        "Pre-treatment average vote share \u2265 5%",
      sample == "At least 10% pre-treatment vote" ~
        "Pre-treatment average vote share \u2265 10%"
    )
  ) |>
  select(sample, selection_criterion, parties, observations)

sample_table
```

```{r table-samples}
# --- FORMAT AND EXPORT APPENDIX TABLE A.1 ---
table_samples <- sample_table |>
  flextable() |>
  set_header_labels(
    sample              = "Sample",
    selection_criterion = "Selection criterion",
    parties             = "No. of parties",
    observations        = "No. of observations"
  ) |>
  theme_booktabs() |>
  align(align = "left", part = "all") |>
  align(j = c("parties", "observations"), align = "center", part = "all") |>  # centred numeric columns
  bold(part = "header") |>
  font(fontname = "Times New Roman", part = "all") |>
  fontsize(size = 10, part = "all") |>
  add_footer_lines(
    c(
      "Note. Observations are party-wave units. All three samples span the same 11 Western European countries.",
      "Source. Own elaboration based on the 1999\u20132024 Chapel Hill Expert Survey trend file (Rovny et al., 2025)."
    )
  ) |>
  italic(part = "footer") |>
  fontsize(size = 9, part = "footer") |>
  set_table_properties(layout = "autofit", width = 1)

table_samples

save_as_docx(table_samples, path = "table3_2_samples.docx")
```

# 3. Descriptive evidence

## 3.1 Treatment timing

```{r treatment-timing-plot}
# --- DESCRIPTIVE: TREATMENT TIMING ---
treatment_timing_plot <- western_country_cohorts |>
  filter(treatment_status != "Treated from first wave") |>
  mutate(
    treatment_label = if_else(
      is.na(treatment_year),
      "Never treated",
      as.character(treatment_year)
    ),
    country_name = forcats::fct_reorder(
      country_name,
      if_else(is.na(treatment_year), 2030, treatment_year)
    )
  ) |>
  ggplot(aes(x = country_name, y = treatment_label, fill = treatment_status)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(
    title = "Treatment timing by country",
    x = NULL,
    y = "First qualifying breakthrough wave"
  ) +
  theme_minimal(base_size = 11)

treatment_timing_plot
```

## 3.2 Aggregate GAL-TAN trajectories

```{r galtan-descriptive-trajectories}
# --- DESCRIPTIVE: GAL-TAN TRAJECTORIES BY TREATMENT STATUS ---
# Raw means by wave; a first visual check of the parallel-trends assumption.
galtan_group_trajectories <- did_panel |>
  mutate(
    comparison_group = case_when(
      treatment_status == "Never treated" ~ "Never treated",
      TRUE ~ "Treated countries"
    )
  ) |>
  group_by(comparison_group, year, wave) |>
  summarise(
    mean_galtan = mean(galtan, na.rm = TRUE),
    n = sum(!is.na(galtan)),
    sd = sd(galtan, na.rm = TRUE),
    se = sd / sqrt(n),
    lower = mean_galtan - 1.96 * se,
    upper = mean_galtan + 1.96 * se,
    .groups = "drop"
  )

last_points_colour <- galtan_group_trajectories |>
  group_by(comparison_group) |>
  slice_max(year, n = 1, with_ties = FALSE) |>
  ungroup()

p_galtan_colour <- ggplot(
  galtan_group_trajectories,
  aes(
    x = year,
    y = mean_galtan,
    colour = comparison_group,
    linetype = comparison_group,
    shape = comparison_group,
    group = comparison_group
  )
) +
  geom_errorbar(
    aes(ymin = lower, ymax = upper),
    width = 0.35,
    linewidth = 0.4
  ) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 2.4, stroke = 0.7) +
  geom_text(
    data = last_points_colour,
    aes(label = comparison_group),
    hjust = 0,
    nudge_x = 1,
    size = 3.2,
    show.legend = FALSE
  ) +
  scale_colour_manual(
    values = c(
      "Never treated" = "#F8766D",
      "Treated countries" = "#00BFC4"
    )
  ) +
  scale_shape_manual(
    values = c(
      "Never treated" = 1,
      "Treated countries" = 16
    )
  ) +
  scale_linetype_manual(
    values = c(
      "Never treated" = "dashed",
      "Treated countries" = "solid"
    )
  ) +
  scale_y_continuous(
    limits = c(0, 10),
    breaks = seq(0, 10, by = 2)
  ) +
  scale_x_continuous(
    breaks = unique(galtan_group_trajectories$year),
    expand = expansion(mult = c(0.02, 0.32))
  ) +
  coord_cartesian(clip = "off") +
  labs(
    x = "CHES wave",
    y = "Mean GAL-TAN position"
  ) +
  theme_classic(base_size = 11) +
  theme(
    legend.position = "none",
    axis.title.y = element_text(
      margin = margin(r = 8)
    ),
    plot.margin = margin(
      t = 5.5,
      r = 45,
      b = 5.5,
      l = 5.5
    )
  )

p_galtan_colour
```

```{r galtan-traject-color}
# --- EXPORT FIGURE (COLOUR) ---
# 5.77 x 4 in at 600 dpi: the figure size used throughout the thesis.
ggsave(
  "figure_galtan_colour.png",
  plot = p_galtan_colour,
  width = 5.77,
  height = 4,
  units = "in",
  dpi = 600
)
```

```{r galtan-traject-bw}
# --- EXPORT FIGURE (BLACK AND WHITE) ---
# Shape and line type carry the grouping so the figure survives printing.
last_points <- galtan_group_trajectories |>
  filter(year == max(year))

ggplot(
  galtan_group_trajectories,
  aes(
    x = year,
    y = mean_galtan,
    linetype = comparison_group,
    shape = comparison_group,
    group = comparison_group
  )
) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.35, colour = "grey40") +
  geom_line(colour = "black", linewidth = 0.6) +
  geom_point(colour = "black", size = 2.4) +
  geom_text(
    data = last_points,
    aes(label = comparison_group),
    hjust = 0,
    nudge_x = 1,
    size = 3.2,
    colour = "black"
  ) +
  scale_shape_manual(values = c("Never treated" = 1, "Treated countries" = 16)) +
  scale_linetype_manual(values = c("Never treated" = "dashed", "Treated countries" = "solid")) +
  scale_y_continuous(limits = c(0, 10), breaks = seq(0, 10, by = 2)) +
  scale_x_continuous(
    breaks = unique(galtan_group_trajectories$year),
    expand = expansion(mult = c(0.02, 0.32))  # more room reserved on the right
  ) +
  coord_cartesian(clip = "off") +  # let the label draw outside the panel if needed
  labs(x = "CHES wave", y = "Mean GAL-TAN position") +
  theme_classic(base_size = 11) +
  theme(
    legend.position = "none",
    plot.margin = margin(t = 5.5, r = 45, b = 5.5, l = 5.5)  # extra physical margin on the right
  )

ggsave(
  "figure_galtan_bw.png",
  width = 5.77, 
  height = 4,
  units = "in",
  dpi = 600
)
```

## 3.3 Descriptive positions on core outcomes

### 3.3.1 Variable operationalization

```{r core-outcomes-defs}
# --- VARIABLE DEFINITIONS (Table 2.2 in the thesis) ---
variable_definitions <- tibble(
  concept = c(
    "GAL-TAN cultural dimension",
    "Immigration policy position",
    "Immigration salience",
    "Economic left-right (placebo)"
  ),
  ches_code = c("galtan", "immigrate_policy", "immigrate_salience", "lrecon"),
  description = c(
    "Party position on democratic and cultural freedoms, ranging from libertarian/postmaterialist to traditional/authoritarian values",
    "Party position on immigration policy",
    "Relative importance of immigration policy in the party's public stance",
    "Party position on economic issues, including taxation, regulation, and the welfare state"
  ),
  scale_range = c(
    "0-10 (0 = Libertarian/Postmaterialist; 5 = Center; 10 = Traditional/Authoritarian)",
    "0-10 (0 = Strongly favors liberal policy; 10 = Strongly favors restrictive policy)",
    "0-10 (0 = Not important at all; 10 = Extremely important)",
    "0-10 (0 = Extreme left; 5 = Center; 10 = Extreme right)"
  )
)

variable_definitions
```

```{r table-vars}
# --- FORMAT AND EXPORT TABLE 2.2 ---
table_variables <- variable_definitions |>
  flextable() |>
  set_header_labels(
    concept     = "Variable concept",
    ches_code   = "CHES code",
    description = "Description",
    scale_range = "Scale range and coding anchors"
  ) |>
  set_table_properties(layout = "autofit", width = 1) |>   # fit to the text width
  font(fontname = "Times New Roman", part = "all") |>
  fontsize(size = 10, part = "header") |>
  fontsize(size = 9,  part = "body") |>
  bold(part = "header") |>
  italic(j = "ches_code", part = "body") |>                # variable code in italics
  valign(valign = "top", part = "body") |>                 # long cells aligned to the top
  border_remove() |>
  hline_top(part = "header",    border = officer::fp_border(width = 1.5)) |>
  hline_bottom(part = "header", border = officer::fp_border(width = 1)) |>
  hline_bottom(part = "body",   border = officer::fp_border(width = 1.5)) |>
  add_footer_lines(
    c(
      "Note. lrecon is included as an economic placebo outcome, not as a substantive dependent variable.",
      "Source. Definitions adapted from the 1999\u20132024 Chapel Hill Expert Survey codebook (Rovny et al., 2025)."
    )
  ) |>
  fontsize(size = 9, part = "footer") |>
  italic(part = "footer")

table_variables

save_as_docx(table_variables, path = "table3_3_variables.docx")
```

### 3.3.2 Summary statistics

```{r summary-statistics}
# --- SUMMARY STATISTICS (Table 2.3 in the thesis) ---
# Build the analytical sample used throughout Section 3.3 as a plain
# data frame with only the four key CHES variables, then preview
# summary statistics with stargazer
summary_data <- did_panel |>
  select(galtan, immigrate_policy, immigrate_salience, lrecon) |>
  as.data.frame()

stargazer(
  summary_data,
  type = "html",
  out  = "table34_summary_stats.html",
  digits = 2,
  covariate.labels = c(
    "GAL-TAN",
    "Immigration policy",
    "Immigration salience",
    "Economic left-right (placebo)"
  ),
  notes = c(
    "Note. N varies across variables because the CHES immigration items (immigration policy and salience) are only available from the 2006 wave onward.",
    "Source. Own elaboration based on the broad analytical sample drawn from the 1999-2024 Chapel Hill Expert Survey trend file (Rovny et al., 2025)."
  ),
  notes.append = FALSE,
  notes.align  = "l"
)
```

```{r core-outcomes-descriptive}
# --- DESCRIPTIVE MEANS BY TREATMENT CONDITION ---
core_outcomes_long <- did_panel |>
  select(
    treatment_status,
    treated,
    galtan,
    immigrate_policy,
    immigrate_salience,
    lrecon
  ) |>
  pivot_longer(
    cols = c(galtan, immigrate_policy, immigrate_salience, lrecon),
    names_to = "outcome",
    values_to = "value"
  ) |>
  filter(!is.na(value)) |>
  mutate(
    period = case_when(
      treatment_status == "Never treated" ~ "Never treated",
      treated == 0 ~ "Pre-treatment",
      treated == 1 ~ "Post-treatment"
    )
  ) |>
  group_by(outcome, period) |>
  summarise(
    n = n(),
    mean = mean(value),
    sd = sd(value),
    se = sd / sqrt(n),
    lower = mean - 1.96 * se,
    upper = mean + 1.96 * se,
    .groups = "drop"
  )

ggplot(core_outcomes_long, aes(x = period, y = mean)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.15) +
  geom_point(size = 2.6) +
  facet_wrap(
    ~ outcome,
    scales = "free_y",
    labeller = as_labeller(c(
      galtan = "GAL-TAN",
      immigrate_policy = "Immigration policy",
      immigrate_salience = "Immigration salience",
      lrecon = "Economic left-right"
    ))
  ) +
  labs(
    title = "Descriptive positions on core outcomes",
    subtitle = "Means with 95% confidence intervals; not causal estimates",
    x = NULL,
    y = "Mean party position"
  ) +
  theme_minimal(base_size = 11)
```

```{r core-outcomes-improved}
# --- SAME AGGREGATION, CHRONOLOGICAL PERIOD ORDER ---
# A shared y-axis range keeps the four outcome panels directly comparable.
core_outcomes_long_bw <- did_panel |>
  select(treatment_status, treated, galtan, immigrate_policy, immigrate_salience, lrecon) |>
  pivot_longer(cols = c(galtan, immigrate_policy, immigrate_salience, lrecon),
               names_to = "outcome", values_to = "value") |>
  filter(!is.na(value)) |>
  mutate(
    period = case_when(
      treatment_status == "Never treated" ~ "Never treated",
      treated == 0 ~ "Pre-treatment",
      treated == 1 ~ "Post-treatment"
    ),
    period = factor(period, levels = c("Never treated", "Pre-treatment", "Post-treatment"))
  ) |>
  group_by(outcome, period) |>
  summarise(n = n(), mean = mean(value), sd = sd(value), se = sd / sqrt(n),
            lower = mean - 1.96 * se, upper = mean + 1.96 * se, .groups = "drop")

y_range <- core_outcomes_long_bw |>
  summarise(y_min = min(lower), y_max = max(upper))

y_limits <- c(floor(y_range$y_min * 2) / 2, ceiling(y_range$y_max * 2) / 2)
y_breaks <- seq(y_limits[1], y_limits[2], by = 0.5)
```

```{r core-outcomes-final}
# --- PANEL BUILDER ---
# One function reused for the four outcomes instead of four near-identical blocks.
make_outcome_panel <- function(data, outcome_name, panel_title, y_label = NULL) {
  data |>
    filter(outcome == outcome_name) |>
    ggplot(aes(x = period, y = mean)) +
    geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.15, colour = "grey40") +
    geom_point(size = 2.4, colour = "black") +
    labs(title = panel_title, x = NULL, y = y_label) +
    scale_y_continuous(limits = y_limits, breaks = y_breaks, labels = scales::label_number(accuracy = 0.1)) +
    theme_classic(base_size = 10) +
    theme(
      plot.title = element_text(size = 10, hjust = 0),
      axis.text.x = element_text(size = 8, angle = 30, hjust = 1),
      axis.title.y = element_text(margin = margin(r = 8))
    )
}

p_galtan    <- make_outcome_panel(core_outcomes_long_bw, "galtan", "GAL-TAN", "Mean party position")
p_immig_pol <- make_outcome_panel(core_outcomes_long_bw, "immigrate_policy", "Immigration policy")
p_immig_sal <- make_outcome_panel(core_outcomes_long_bw, "immigrate_salience", "Immigration salience", "Mean party position")
p_lrecon    <- make_outcome_panel(core_outcomes_long_bw, "lrecon", "Economic left-right")

(p_galtan + p_immig_pol) / (p_immig_sal + p_lrecon) +
  plot_annotation(tag_levels = "A")
```

```{r core-outcomes-export}
# --- EXPORT FIGURE ---
ggsave(
  "figure_core_outcomes_bw.png",
  width = 5.77,
  height = 5.5,
  units = "in",
  dpi = 600
)
```

# 4. Empirical strategy

## 4.1 TWFE benchmark and Sun-Abraham event study

```{r twfe-and-sunab}
# --- TWFE BENCHMARK AND SUN-ABRAHAM EVENT STUDY ---
# TWFE is reported as a benchmark only: with staggered timing and heterogeneous
# effects it is biased (negative weights, forbidden comparisons).
# sunab() needs a finite cohort id, so never-treated units are coded 1000.
twfe_galtan <- feols(
  galtan ~ treated | party_id + wave,
  data = did_panel,
  cluster = ~country
)

did_panel_sa <- did_panel |>
  mutate(cohort_sa = if_else(g_wave == 0, 1000L, g_wave))

sunab_galtan <- feols(
  galtan ~ sunab(cohort_sa, wave) | party_id + wave,
  data = did_panel_sa,
  cluster = ~country
)

summary(twfe_galtan)
summary(sunab_galtan, agg = "att")
```

```{r sunab-galtan-plot}
# --- SUN-ABRAHAM EVENT-STUDY PLOT ---
iplot(
  sunab_galtan,
  ref.line = -0.5,
  xlab = "CHES waves relative to treatment",
  main = "Sun-Abraham event study: GAL-TAN"
)
```

# 5. Main results

## 5.1 GAL-TAN

```{r callaway-santanna-galtan}
# --- MAIN ESTIMATE: GAL-TAN (CALLAWAY-SANT'ANNA) ---
# control_group = 'notyettreated' enlarges the comparison pool, which matters
# with this few country clusters.
# clustervars = 'country' because treatment is assigned at country level.
# allow_unbalanced_panel = TRUE: parties enter and leave CHES waves.
# The fixed seed makes the multiplier-bootstrap SEs exactly reproducible.
set.seed(1234) # Fixed seed ensures reproducible bootstrap SEs/CIs (Callaway & Sant'Anna estimator)

cs_galtan <- att_gt(
  yname = "galtan",
  tname = "wave",
  idname = "party_id",
  gname = "g_wave",
  data = did_panel,
  control_group = "notyettreated",
  clustervars = "country",
  allow_unbalanced_panel = TRUE
)

cs_galtan_simple <- aggte(
  cs_galtan,
  type = "simple",
  na.rm = TRUE
)

summary(cs_galtan_simple)
```

## 5.2 Immigration policy

```{r immigration-panel-and-estimate}
# --- MAIN ESTIMATE: IMMIGRATION POLICY ---
# Restricted to 2006 onwards (item availability); the Netherlands is dropped
# because of consecutive missing values on the immigration items.
did_panel_immigration <- did_panel |>
  filter(
    year >= 2006,
    !is.na(immigrate_policy),
    country_name != "Netherlands"
  )

set.seed(1234)

cs_immigration <- att_gt(
  yname = "immigrate_policy",
  tname = "wave",
  idname = "party_id",
  gname = "g_wave",
  data = did_panel_immigration,
  control_group = "notyettreated",
  clustervars = "country",
  allow_unbalanced_panel = TRUE
)

cs_immigration_simple <- aggte(
  cs_immigration,
  type = "simple",
  na.rm = TRUE
)

summary(cs_immigration_simple)
```

## 5.3 Immigration issue salience

```{r immigration-salience-estimate}
# --- MAIN ESTIMATE: IMMIGRATION SALIENCE ---
did_panel_immigration_salience <- did_panel |>
  filter(
    year >= 2006,
    !is.na(immigrate_salience),
    country_name != "Netherlands"
  )

set.seed(1234)

cs_immigration_salience <- att_gt(
  yname = "immigrate_salience",
  tname = "wave",
  idname = "party_id",
  gname = "g_wave",
  data = did_panel_immigration_salience,
  control_group = "notyettreated",
  clustervars = "country",
  allow_unbalanced_panel = TRUE
)

cs_immigration_salience_simple <- aggte(
  cs_immigration_salience,
  type = "simple",
  na.rm = TRUE
)

summary(cs_immigration_salience_simple)
```

## 5.4 Main-results table and forest plot

```{r main-results-table}
# --- MAIN RESULTS TABLE DATA ---
main_results <- bind_rows(
  tibble(
    sample = "Broad sample",
    outcome = "GAL-TAN",
    att = cs_galtan_simple$overall.att,
    se = cs_galtan_simple$overall.se
  ),
  tibble(
    sample = "Broad sample",
    outcome = "Immigration policy",
    att = cs_immigration_simple$overall.att,
    se = cs_immigration_simple$overall.se
  ),
  tibble(
    sample = "Broad sample",
    outcome = "Immigration salience",
    att = cs_immigration_salience_simple$overall.att,
    se = cs_immigration_salience_simple$overall.se
  )
) |>
  mutate(
    lower = att - 1.96 * se,
    upper = att + 1.96 * se
  )

main_results
```

```{r main-results-forest-plot}
# --- MAIN RESULTS FOREST PLOT ---
ggplot(
  main_results,
  aes(
    x = att,
    y = forcats::fct_reorder(outcome, att)
  )
) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_errorbar(
  aes(xmin = lower, xmax = upper),
  width = 0.15,
  orientation = "y"
) +
  geom_point(size = 2.8) +
  labs(
    title = "Average treatment effects on core outcomes",
    subtitle = "Callaway-Sant'Anna estimates, broad sample, 95% confidence intervals",
    x = "Average treatment effect on the treated",
    y = NULL
  ) +
  theme_minimal(base_size = 11)
```

# 6. Diagnostics and robustness

## 6.1 Dynamic effects and pre-treatment trajectories

```{r dynamic-effects}
# --- DYNAMIC (EVENT-TIME) AGGREGATION ---
# cband = FALSE returns pointwise confidence intervals, which is what the
# thesis figures and their notes report.
set.seed(1234)

cs_galtan_dynamic <- aggte(
  cs_galtan,
  type = "dynamic",
  na.rm = TRUE,
  cband = FALSE
)

set.seed(1234)

cs_immigration_dynamic <- aggte(
  cs_immigration,
  type = "dynamic",
  na.rm = TRUE,
  cband = FALSE
)

summary(cs_galtan_dynamic)
summary(cs_immigration_dynamic)
```

```{r dynamic-galtan-plot}
# --- DYNAMIC EFFECTS PLOT: GAL-TAN (Appendix Figure B.1) ---
p_galtan_dynamic <- ggdid(
  cs_galtan_dynamic,
  xlab = "CHES waves relative to treatment",
  ylab = "Average treatment effect",
  title = " ",
  legend = TRUE,
  ref_line = 0,
  theming = TRUE
) +
  geom_vline(
    xintercept = -0.5,
    linetype = "dotted",
    linewidth = 0.4,
    colour = "grey50"
  ) +
  labs(
    title = NULL,
    subtitle = NULL,
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    axis.title.y = element_text(
      margin = margin(r = 8)
    ),
    plot.margin = margin(
      t = 5.5,
      r = 5.5,
      b = 5.5,
      l = 5.5
    )
  )

p_galtan_dynamic
```

```{r dynamic-galtan-export}
# --- EXPORT APPENDIX FIGURE B.1 ---
ggsave(
  "figure_dynamic_galtan_colour.png",
  plot = p_galtan_dynamic,
  width = 5.77,
  height = 4,
  units = "in",
  dpi = 600,
  bg = "white"
)
```

```{r dynamic-immigration-table-data}
# --- WAVE-BY-WAVE ESTIMATES (Appendix Table B.2) ---
immigration_dynamic_results <- tibble(
  event_time = cs_immigration_dynamic$egt,
  att = cs_immigration_dynamic$att.egt,
  se = cs_immigration_dynamic$se.egt
) |>
  mutate(
    period = if_else(
      event_time < 0,
      "Pre-treatment",
      "Post-treatment"
    ),
    lower = att - 1.96 * se,
    upper = att + 1.96 * se
  ) |>
  select(
    event_time,
    period,
    att,
    se,
    lower,
    upper
  ) |>
  arrange(event_time)

immigration_dynamic_results
```

```{r dynamic-immigration-table-format}
# --- FORMAT APPENDIX TABLE B.2 ---
immigration_dynamic_fmt <- immigration_dynamic_results |>
  mutate(
    stars = if_else(lower > 0 | upper < 0, "*", ""),
    event_time = case_when(
      event_time > 0 ~ paste0("+", event_time),
      TRUE ~ as.character(event_time)
    ),
    `ATT` = paste0(sprintf("%.3f", att), stars),
    `SE` = sprintf("%.3f", se),
    `95% CI` = sprintf("[%.3f, %.3f]", lower, upper)
  ) |>
  select(
    `Event time` = event_time,
    `Period` = period,
    `ATT`,
    `SE`,
    `95% CI`
  )

immigration_dynamic_fmt
```

```{r dynamic-immigration-table-export}
# --- EXPORT APPENDIX TABLE B.2 ---
stargazer(
  as.data.frame(immigration_dynamic_fmt),
  type = "html",
  summary = FALSE,
  rownames = FALSE,
  notes = c(
    "<em>Note.</em> ATT = average treatment effect on the treated; SE = standard error; CI = confidence interval. Event time 0 denotes the first CHES wave in which treatment is observed. Negative and positive values denote waves before and after treatment, respectively. * indicates that the 95% pointwise confidence interval excludes zero.",
    "<em>Source.</em> Own elaboration based on the 1999–2024 Chapel Hill Expert Survey trend file (Rovny et al., 2025)."
  ),
  notes.append = FALSE,
  notes.align = "l",
  out = "tableA1_dynamic_immigration.html"
)
```

```{r dynamic-immigration-plot}
# --- DYNAMIC EFFECTS PLOT: IMMIGRATION (Appendix Figure B.2) ---
p_immigration_dynamic <- ggdid(
  cs_immigration_dynamic,
  xlab = "CHES waves relative to treatment",
  ylab = "Average treatment effect",
  title = " ",
  legend = TRUE,
  ref_line = 0,
  theming = TRUE
) +
  geom_vline(
    xintercept = -0.5,
    linetype = "dotted",
    linewidth = 0.4,
    colour = "grey50"
  ) +
  labs(
    title = NULL,
    subtitle = NULL,
    colour = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    axis.title.y = element_text(
      margin = margin(r = 8)
    ),
    plot.margin = margin(
      t = 5.5,
      r = 5.5,
      b = 5.5,
      l = 5.5
    )
  )

p_immigration_dynamic
```

```{r dynamic-immigration-export}
# --- EXPORT APPENDIX FIGURE B.2 ---
ggsave(
  "figure_dynamic_immigration_colour.png",
  plot = p_immigration_dynamic,
  width = 5.77,
  height = 4,
  units = "in",
  dpi = 600,
  bg = "white"
)
```

## 6.2 Comparative summary table on core mainstream positioning

```{r table-comparative-data}
# --- COMPARATIVE ESTIMATES TABLE DATA (Table 3.1) ---
table_comparative_core <- bind_rows(
  tibble(
    outcome = "GAL-TAN (galtan)",
    estimator = "TWFE Benchmark (OLS)",
    att = coef(twfe_galtan)["treated"],
    se = se(twfe_galtan)["treated"]
  ),
  tibble(
    outcome = "GAL-TAN (galtan)",
    estimator = "Sun-Abraham (IW)",
    att = summary(sunab_galtan, agg = "att")$coeftable[1, "Estimate"],
    se = summary(sunab_galtan, agg = "att")$coeftable[1, "Std. Error"]
  ),
  tibble(
    outcome = "GAL-TAN (galtan)",
    estimator = "Callaway-Sant'Anna (Simple aggregation)",
    att = cs_galtan_simple$overall.att,
    se = cs_galtan_simple$overall.se
  ),
  tibble(
    outcome = "GAL-TAN (galtan)",
    estimator = "Callaway-Sant'Anna (Dynamic aggregation)",
    att = cs_galtan_dynamic$overall.att,
    se = cs_galtan_dynamic$overall.se
  ),
  tibble(
    outcome = "Immigration position (immigrate_policy)",
    estimator = "Callaway-Sant'Anna (Simple aggregation)",
    att = cs_immigration_simple$overall.att,
    se = cs_immigration_simple$overall.se
  ),
  tibble(
    outcome = "Immigration policy (immigrate_policy)",
    estimator = "Callaway-Sant'Anna (Dynamic aggregation)",
    att = cs_immigration_dynamic$overall.att,
    se = cs_immigration_dynamic$overall.se
  ),
  tibble(
    outcome = "Immigration salience (immigrate_salience)",
    estimator = "Callaway-Sant'Anna (Simple aggregation)",
    att = cs_immigration_salience_simple$overall.att,
    se = cs_immigration_salience_simple$overall.se
  )
) |>
  mutate(
    lower = att - 1.96 * se,
    upper = att + 1.96 * se
  )

table_comparative_core
```

```{r table-comparative-format}
# --- FORMAT TABLE 3.1 ---
table_comparative_fmt <- table_comparative_core |>
  mutate(
    stars = if_else(lower > 0 | upper < 0, "*", ""),
    `ATT` = paste0(sprintf("%.3f", att), stars),
    `SE` = sprintf("%.3f", se),
    `95% CI` = sprintf("[%.3f, %.3f]", lower, upper),
    outcome = if_else(duplicated(outcome), "", outcome)
  ) |>
  select(
  `CHES outcome` = outcome,
  `Estimator / aggregation` = estimator,
  `ATT`,
  `SE`,
  `95% CI`
)

table_comparative_fmt
```

```{r table-comparative-export}
# --- EXPORT TABLE 3.1 ---
stargazer(
  as.data.frame(table_comparative_fmt),
  type = "html",
  summary = FALSE,
  rownames = FALSE,
  notes = c(
  "<em>Note.</em> ATT = average treatment effect on the treated; SE = standard error; CI = confidence interval; TWFE = two-way fixed effects; IW = interaction-weighted event study. * indicates that the 95% confidence interval excludes zero. Callaway-Sant'Anna group-time ATTs are estimated using the doubly robust method; simple and dynamic refer to alternative aggregation schemes.",
  "<em>Source.</em> Own elaboration based on the 1999–2024 Chapel Hill Expert Survey trend file (Rovny et al., 2025)."
),
  notes.append = FALSE,
  notes.align = "l",
  out = "table6_2_comparative.html"
)
```

## 6.3 Alternative comparison group

```{r never-treated-robustness}
# --- ROBUSTNESS: NEVER-TREATED COMPARISON GROUP ---
# Same specification as the main model with control_group = 'nevertreated',
# to check that the flat cultural result is not an artefact of the larger
# not-yet-treated control pool.
cs_galtan_never <- att_gt(
  yname = "galtan",
  tname = "wave",
  idname = "party_id",
  gname = "g_wave",
  data = did_panel,
  control_group = "nevertreated",
  clustervars = "country",
  allow_unbalanced_panel = TRUE
)

cs_galtan_never_simple <- aggte(
  cs_galtan_never, type = "simple", na.rm = TRUE
)

summary(cs_galtan_never_simple)
```

## 6.4 Electoral-size thresholds

```{r threshold-5-percent-main-outcomes}
# --- ROBUSTNESS: 5% PRE-TREATMENT VOTE SAMPLE (Appendix Table B.1) ---
did_panel_5_immigration <- did_panel_5 |>
  filter(
    year >= 2006,
    !is.na(immigrate_policy),
    country_name != "Netherlands"
  )

cs_galtan_5 <- att_gt(
  yname = "galtan",
  tname = "wave",
  idname = "party_id",
  gname = "g_wave",
  data = did_panel_5,
  control_group = "notyettreated",
  clustervars = "country",
  allow_unbalanced_panel = TRUE
)

cs_galtan_5_simple <- aggte(cs_galtan_5, type = "simple", na.rm = TRUE)

cs_immigration_5 <- att_gt(
  yname = "immigrate_policy",
  tname = "wave",
  idname = "party_id",
  gname = "g_wave",
  data = did_panel_5_immigration,
  control_group = "notyettreated",
  clustervars = "country",
  allow_unbalanced_panel = TRUE
)

cs_immigration_5_simple <- aggte(
  cs_immigration_5,
  type = "simple",
  na.rm = TRUE
)

summary(cs_galtan_5_simple)
summary(cs_immigration_5_simple)
```

```{r threshold-10-percent-galtan}
# --- ROBUSTNESS: 10% PRE-TREATMENT VOTE SAMPLE (Appendix Table B.1) ---
cs_galtan_10 <- att_gt(
  yname = "galtan",
  tname = "wave",
  idname = "party_id",
  gname = "g_wave",
  data = did_panel_10,
  control_group = "notyettreated",
  clustervars = "country",
  allow_unbalanced_panel = TRUE
)

cs_galtan_10_simple <- aggte(cs_galtan_10, type = "simple", na.rm = TRUE)

summary(cs_galtan_10_simple)
```

## 6.5 Outcome specificity

```{r outcome-specificity}
# --- OUTCOME SPECIFICITY AND ECONOMIC PLACEBO ---
# multiculturalism: alternative cultural indicator.
# lrecon: economic placebo. A significant effect here would indicate that
# breakthrough timing coincides with broader programmatic adjustment rather
# than an issue-specific cultural response.
did_panel_multiculturalism <- did_panel |>
  filter(
    year >= 2006,
    !is.na(multiculturalism),
    country_name != "Netherlands"
  )

cs_multiculturalism <- att_gt(
  yname = "multiculturalism",
  tname = "wave",
  idname = "party_id",
  gname = "g_wave",
  data = did_panel_multiculturalism,
  control_group = "notyettreated",
  clustervars = "country",
  allow_unbalanced_panel = TRUE
)

cs_multiculturalism_simple <- aggte(
  cs_multiculturalism,
  type = "simple",
  na.rm = TRUE
)

cs_lrecon <- att_gt(
  yname = "lrecon",
  tname = "wave",
  idname = "party_id",
  gname = "g_wave",
  data = did_panel,
  control_group = "notyettreated",
  clustervars = "country",
  allow_unbalanced_panel = TRUE
)

cs_lrecon_simple <- aggte(cs_lrecon, type = "simple", na.rm = TRUE)

cs_lrecon_5 <- att_gt(
  yname = "lrecon",
  tname = "wave",
  idname = "party_id",
  gname = "g_wave",
  data = did_panel_5,
  control_group = "notyettreated",
  clustervars = "country",
  allow_unbalanced_panel = TRUE
)

cs_lrecon_5_simple <- aggte(cs_lrecon_5, type = "simple", na.rm = TRUE)

outcome_specificity_results <- bind_rows(
  tibble(
    sample = "Broad sample",
    outcome = "Multiculturalism",
    att = cs_multiculturalism_simple$overall.att,
    se = cs_multiculturalism_simple$overall.se
  ),
  tibble(
    sample = "Broad sample",
    outcome = "Economic left-right",
    att = cs_lrecon_simple$overall.att,
    se = cs_lrecon_simple$overall.se
  ),
  tibble(
    sample = "5% threshold",
    outcome = "Economic left-right",
    att = cs_lrecon_5_simple$overall.att,
    se = cs_lrecon_5_simple$overall.se
  )
) |>
  mutate(
    lower = att - 1.96 * se,
    upper = att + 1.96 * se
  )

outcome_specificity_results
```

```{r robustness-results-data}
# --- ROBUSTNESS AND PLACEBO RESULTS: DATA (Appendix Table B.1) ---
# Collects the estimates produced in 6.3-6.5 into a single tidy table.
robustness_placebo_results <- bind_rows(
  tibble(
    category = "Alternative control group",
    specification = "Never-treated controls only",
    outcome = "GAL-TAN (galtan)",
    att = cs_galtan_never_simple$overall.att,
    se = cs_galtan_never_simple$overall.se
  ),
  tibble(
    category = "Electoral-size thresholds",
    specification = "5% pre-treatment vote sample",
    outcome = "GAL-TAN (galtan)",
    att = cs_galtan_5_simple$overall.att,
    se = cs_galtan_5_simple$overall.se
  ),
  tibble(
    category = "Electoral-size thresholds",
    specification = "10% pre-treatment vote sample",
    outcome = "GAL-TAN (galtan)",
    att = cs_galtan_10_simple$overall.att,
    se = cs_galtan_10_simple$overall.se
  ),
  tibble(
    category = "Electoral-size thresholds",
    specification = "5% pre-treatment vote sample",
    outcome = "Immigration policy (immigrate_policy)",
    att = cs_immigration_5_simple$overall.att,
    se = cs_immigration_5_simple$overall.se
  ),
  tibble(
    category = "Outcome-specificity check",
    specification = "Broad sample",
    outcome = "Multiculturalism (multiculturalism)",
    att = cs_multiculturalism_simple$overall.att,
    se = cs_multiculturalism_simple$overall.se
  ),
  tibble(
    category = "Placebo-dimension check",
    specification = "Broad sample",
    outcome = "Economic left-right (lrecon)",
    att = cs_lrecon_simple$overall.att,
    se = cs_lrecon_simple$overall.se
  ),
  tibble(
    category = "Placebo-dimension check",
    specification = "5% pre-treatment vote sample",
    outcome = "Economic left-right (lrecon)",
    att = cs_lrecon_5_simple$overall.att,
    se = cs_lrecon_5_simple$overall.se
  )
) |>
  mutate(
    lower = att - 1.96 * se,
    upper = att + 1.96 * se
  )

robustness_placebo_results
```

```{r robustness-results-format}
# --- FORMAT APPENDIX TABLE B.1 ---
robustness_placebo_fmt <- robustness_placebo_results |>
  mutate(
    stars = if_else(lower > 0 | upper < 0, "*", ""),
    category_display = if_else(
      duplicated(category),
      "",
      category
    ),
    `ATT` = paste0(sprintf("%.3f", att), stars),
    `SE` = sprintf("%.3f", se),
    `95% CI` = sprintf("[%.3f, %.3f]", lower, upper)
  ) |>
  select(
    `Robustness category` = category_display,
    `Sample / specification` = specification,
    `Dependent variable (CHES)` = outcome,
    `ATT`,
    `SE`,
    `95% CI`
  )

robustness_placebo_fmt
```

```{r robustness-results-export}
# --- EXPORT APPENDIX TABLE B.1 ---
stargazer(
  as.data.frame(robustness_placebo_fmt),
  type = "html",
  summary = FALSE,
  rownames = FALSE,
  notes = c(
    "<em>Note.</em> Entries report doubly robust simple aggregate average treatment effects on the treated (ATTs) estimated using the Callaway and Sant’Anna staggered difference-in-differences estimator. Standard errors are clustered at the country level. CI = confidence interval. * indicates that the 95% confidence interval excludes zero. The alternative-control specification uses only never-treated countries; all other specifications use not-yet-treated units as the comparison group. Immigration policy and multiculturalism estimates use outcome-available observations from 2006 onward and exclude the Netherlands.",
    "<em>Source.</em> Own elaboration based on the 1999–2024 Chapel Hill Expert Survey trend file (Rovny et al., 2025)."
  ),
  notes.append = FALSE,
  notes.align = "l",
  out = "table4_2_robustness.html"
)
```

## 6.6 Summary table of robustness checks

```{r robustness-summary-superseded, eval=FALSE}
# --- SUPERSEDED DRAFT TABLE (not run) ---
# Kept for the record only. It calls format_att_table() on an object
# (robustness_results) that is never created in this script, and the table it
# was meant to produce is the one built and exported just above. eval=FALSE so
# the document knits; delete the chunk if it is no longer wanted.
format_att_table(
  robustness_results,
  caption = "Table X. Callaway-Sant’Anna estimates and robustness checks"
)
```

# 7. Heterogeneity by party family

## 7.1 Mainstream right versus mainstream left

```{r party-family-heterogeneity}
# --- HETEROGENEITY BY PARTY FAMILY (H3) ---
# Mainstream right (conservative and Christian-democratic families) against
# mainstream left (socialist family), on the 5% pre-treatment vote sample.
estimate_galtan_att <- function(data) {
  att_gt(
    yname = "galtan",
    tname = "wave",
    idname = "party_id",
    gname = "g_wave",
    data = data,
    control_group = "notyettreated",
    clustervars = "country",
    allow_unbalanced_panel = TRUE
  )
}

cs_galtan_right_5 <- estimate_galtan_att(
  did_panel_5 |>
    filter(family %in% c(2, 4))
)

cs_galtan_left_5 <- estimate_galtan_att(
  did_panel_5 |>
    filter(family == 5)
)

galtan_right_left_5 <- bind_rows(
  tibble(
    group = "Mainstream right",
    att = aggte(cs_galtan_right_5, type = "simple", na.rm = TRUE)$overall.att,
    se = aggte(cs_galtan_right_5, type = "simple", na.rm = TRUE)$overall.se
  ),
  tibble(
    group = "Mainstream left",
    att = aggte(cs_galtan_left_5, type = "simple", na.rm = TRUE)$overall.att,
    se = aggte(cs_galtan_left_5, type = "simple", na.rm = TRUE)$overall.se
  )
) |>
  mutate(
    lower = att - 1.96 * se,
    upper = att + 1.96 * se
  )

galtan_right_left_5
```

# 8. Optional diagnostics

```{r optional-party-country-trajectories, eval=FALSE}
# --- OPTIONAL DIAGNOSTICS (not run: eval=FALSE) ---
```
